---
title: "NINESIM Discrete Event Simulation Plots"
author: "Jan Irvahn"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r read in data,echo=FALSE}
library(ggplot2)

system("python qdriver.py > out.txt")
system("grep \"\\[\" out.txt > timestamps.csv")
system("sed -i -e 's/\\[//g' timestamps.csv") 
system("sed -i -e 's/\\]//g' timestamps.csv")

st<-read.csv("timestamps.csv",header=FALSE)
names(st)<-c("arrival","leave_arrival","enter_primary_queue",
             "leave_primary_queue/start_primary_screening",
             "end_primary_screening_if_secondary_bound","enter_secondary_queue",
             "leave_secondary_queue/start_secondary_screening",
             "leave_secondary_screening_OR_leave_primary_for_exit",
             "exit")

primary_processing_time<-function(ol) {
  if(ol[5]==0) return(ol[8]-ol[4])
  if(ol[5]!=0) return(ol[5]-ol[4])
}
ppt<-apply(st,1,primary_processing_time)

mdf<-data.frame(st)
secondary<-rep("no",nrow(mdf))
secondary[mdf$enter_secondary_queue!=0]<-"yes"
ndf<-cbind(mdf,secondary,ppt)

line_mql<-system("grep \"QMAX\" qdriver.py",intern=TRUE)
snas<-unlist(strsplit(unlist(strsplit(line_mql,split="\\["))[2],","))
pl1qm<-as.numeric(snas[3])
pl2qm<-as.numeric(snas[4])
pl3qm<-as.numeric(snas[5])
sqm<-as.numeric(snas[6])

line_n<-system("grep \"NUM_VEHICLES\" qdriver.py",intern=TRUE)
n<-as.numeric(unlist(strsplit(unlist(strsplit(line_n,split=":"))[2],","))[1])

line_p<-system("grep \"probability of being sent to\" qdriver.py",intern=TRUE)
p<-as.numeric(unlist(strsplit(unlist(strsplit(line_p,split=":"))[2],","))[1])

line_mtba<-system("grep \"mean time between arrivals\" qdriver.py",intern=TRUE)
mean_arrival<-as.numeric(unlist(strsplit(unlist(strsplit(line_mtba,split=":"))[2],","))[1])

line_mtip<-system("grep \"mean time in primary\" qdriver.py",intern=TRUE)
mean_primary<-as.numeric(unlist(strsplit(unlist(strsplit(line_mtip,split=":"))[2],","))[1])

line_mtis<-system("grep \"mean time in secondary\" qdriver.py",intern=TRUE)
mean_secondary<-as.numeric(unlist(strsplit(unlist(strsplit(line_mtis,split=":"))[2],","))[1])

line_tt<-system("grep \"time to drive to a queue\" qdriver.py",intern=TRUE)
transit_time<-as.numeric(unlist(strsplit(unlist(strsplit(line_tt,split=":"))[2],"#"))[1])

options(scipen=1, digits=2)
```

# Introduction

PNNL has started writing python code for a discrete event simulation of the queueing model shown in Figure \ref{EG} using the SimianPie package that was developed at LANL. 

## Parameter Settings

The queueing model has 3 lanes at primary screening. The time between arrivals was exponentially distributed with a mean of `r mean_arrival` minutes. Primary screening times were exponentially distributed with a mean of `r mean_primary` minutes. Secondary screening times were exponentially distributed with a mean of `r mean_secondary` minutes. Transit times to get to the primary queues, the secondary queue, and the exit were `r transit_time` minutes. Maximum queue length for the three primary queues were `r pl1qm`, `r pl2qm`, and `r pl3qm`. The maximum queue length for secondary was `r sqm`. The probability of a vehicle being sent to secondary was `r p`.  The simulation described here ran `r signif(n,5)` vehicles. 

\begin{figure}[!h]
\includegraphics{eventgraph.pdf}
\caption{Event graph for NINESIM queueing work. \label{EG}}
\end{figure}

\clearpage

# Plots of Simulation Results

```{r arrival queue wait times,echo=FALSE,fig.cap='Time spent in the arrival queue as a function of arrival time after the start of the simulation.'}
ggplot(ndf,aes(x=arrival,y=leave_arrival-arrival))+
  geom_point()+
  theme_bw()+ 
  theme(panel.border=element_blank(),panel.grid.major=element_blank(),panel.grid.minor=element_blank(),axis.line=element_line(colour = "black"))+
  labs(title="Arrival Queue Wait Times vs. Arrival Time")+
  ylab("Arrival Queue Wait Times (minutes)")+
  xlab("Arrival Time (minutes)")+
  theme(plot.title = element_text(hjust = 0.5))
```

```{r hist of arrival queue wait times,echo=FALSE,fig.cap='Histogram of the times spent in the arrival queue.'}
ggplot(ndf,aes(x=leave_arrival-arrival))+
  geom_histogram(color="black",bins=30)+
  theme_bw()+ 
  theme(panel.border=element_blank(),panel.grid.major=element_blank(),panel.grid.minor=element_blank(),axis.line=element_line(colour = "black"))+
  labs(title="Arrival Queue Wait Times")+
  xlab("Minutes")+
  theme(plot.title = element_text(hjust = 0.5))
```

```{r hist of primary queue wait times,echo=FALSE,fig.cap='Histogram of the times spent in the queue for primary screening.'}
ggplot(ndf,aes(x=leave_primary_queue.start_primary_screening-enter_primary_queue))+
  geom_histogram(color="black",bins=30)+
  theme_bw()+ 
  theme(panel.border=element_blank(),panel.grid.major=element_blank(),panel.grid.minor=element_blank(),axis.line=element_line(colour = "black"))+
  labs(title="Primary Queue Wait Times")+
  xlab("Minutes")+
  theme(plot.title = element_text(hjust = 0.5))
```

```{r hist of primary screening times,echo=FALSE,fig.cap='Histogram of the times spent in primary screening. This includes time waiting for space in the secondary queue.'}
ggplot(ndf,aes(x=ppt))+
  geom_histogram(color="black",bins=30)+
  theme_bw()+ 
  theme(panel.border=element_blank(),panel.grid.major=element_blank(),panel.grid.minor=element_blank(),axis.line=element_line(colour = "black"))+
  labs(title="Primary Screening Times")+
  xlab("Minutes")+
  theme(plot.title = element_text(hjust = 0.5))
```

```{r hist of secondary queue wait times,echo=FALSE,fig.cap='Histogram of the times spent in the queue for secondary screening.'}
tdf<-ndf[ndf$enter_secondary_queue!=0,]
ggplot(tdf,aes(x=leave_secondary_queue.start_secondary_screening-enter_secondary_queue))+
  geom_histogram(color="black",bins=30)+
  theme_bw()+ 
  theme(panel.border=element_blank(),panel.grid.major=element_blank(),panel.grid.minor=element_blank(),axis.line=element_line(colour = "black"))+
  labs(title="Secondary Queue Wait Times")+
  xlab("Minutes")+
  theme(plot.title = element_text(hjust = 0.5))
```

```{r hist of secondary screening times,echo=FALSE,fig.cap='Histogram of the times spent in secondary screening.'}
ggplot(tdf,aes(x=leave_secondary_screening_OR_leave_primary_for_exit-
                 leave_secondary_queue.start_secondary_screening))+
  geom_histogram(color="black",bins=30)+
  theme_bw()+ 
  theme(panel.border=element_blank(),panel.grid.major=element_blank(),panel.grid.minor=element_blank(),axis.line=element_line(colour = "black"))+
  labs(title="Secondary Screening Times")+
  xlab("Minutes")+
  theme(plot.title = element_text(hjust = 0.5))
```

```{r total time by arrival time,echo=FALSE,fig.cap='Total time in system as a function of arrival time.'}
ggplot(ndf,aes(x=arrival,y=exit-arrival,color=secondary))+
  geom_point()+
  theme_bw()+ 
  theme(panel.border=element_blank(),panel.grid.major=element_blank(),panel.grid.minor=element_blank(),axis.line=element_line(colour = "black"))+
  labs(title="Total Time in System vs. Arrival Time")+
  ylab("Total Time in System (minutes)")+
  xlab("Arrival Time (minutes)")+
  theme(plot.title = element_text(hjust = 0.5))
```

```{r total time in system,echo=FALSE,fig.cap='Histogram of total time in system.'}
ggplot(ndf,aes(x=exit-arrival,color=secondary,fill=secondary))+
  geom_histogram(alpha=0.5,bins=30)+
  theme_bw()+ 
  theme(panel.border=element_blank(),panel.grid.major=element_blank(),panel.grid.minor=element_blank(),axis.line=element_line(colour = "black"))+
  labs(title="Total Time in System")+
  xlab("Total Time in System (minutes)")+
  theme(plot.title = element_text(hjust = 0.5))
```

```{r queue states set up,echo=FALSE}
options(stringsAsFactors = FALSE)
system("grep \"queue length\" out.txt > queuelengths.csv")
ql<-read.csv("queuelengths.csv",header=FALSE)
extract_queue_lengths<-function(x){
 is<-unlist(strsplit(x,split=":"))
 time<-as.numeric(is[1])
 queue_length<-as.numeric(is[3])
 ald<-is[2]
 queue_label<-as.numeric(unlist(strsplit(unlist(strsplit(ald,split="\\("))[2],"\\)"))[1])
 rdf<-data.frame(time=time,queue_label=queue_label,queue_length=queue_length)
 return(rdf)
}
library(plyr)
hat<-rbind.fill(apply(ql,1,extract_queue_lengths))
# arrival queue length versus time
qol<-hat[hat$queue_label==1,]
qol<-qol[order(qol$time),]
p0<-ggplot(qol,aes(x=time,y=queue_length))+
  geom_step()+
  theme_bw()+ 
  theme(panel.border=element_blank(),panel.grid.major=element_blank(),panel.grid.minor=element_blank(),axis.line=element_line(colour = "black"))+
  labs(title="Arrival Queue Length vs. Arrival Time")+
  ylab("Arrival Queue Length")+
  xlab("Time (minutes after simulation start)")+
  theme(plot.title = element_text(hjust = 0.5))
# lane 1 queue length versus time
qtwol<-hat[hat$queue_label==2,]
qtwol<-qtwol[order(qtwol$time),]
p1<-ggplot(qtwol,aes(x=time,y=queue_length))+
  geom_step()+
  theme_bw()+ 
  theme(panel.border=element_blank(),panel.grid.major=element_blank(),panel.grid.minor=element_blank(),axis.line=element_line(colour = "black"))+
  labs(title="Primary Lane 1 Queue Length")+
  ylab("Primary Lane 1 Queue Length")+
  xlab("Time (minutes after simulation start)")+
  theme(plot.title = element_text(hjust = 0.5))
# lane 2 queue length versus time
qthreel<-hat[hat$queue_label==3,]
qthreel<-qthreel[order(qthreel$time),]
p2<-ggplot(qthreel,aes(x=time,y=queue_length))+
  geom_step()+
  theme_bw()+ 
  theme(panel.border=element_blank(),panel.grid.major=element_blank(),panel.grid.minor=element_blank(),axis.line=element_line(colour = "black"))+
  labs(title="Primary Lane 2 Queue Length")+
  ylab("Primary Lane 2 Queue Length")+
  xlab("Time (minutes after simulation start)")+
  theme(plot.title = element_text(hjust = 0.5))
# lane 3 queue length versus time
qfourl<-hat[hat$queue_label==4,]
qfourl<-qfourl[order(qfourl$time),]
p3<-ggplot(qfourl,aes(x=time,y=queue_length))+
  geom_step()+
  theme_bw()+ 
  theme(panel.border=element_blank(),panel.grid.major=element_blank(),panel.grid.minor=element_blank(),axis.line=element_line(colour = "black"))+
  labs(title="Primary Lane 3 Queue Length")+
  ylab("Primary Lane 3 Queue Length")+
  xlab("Time (minutes after simulation start)")+
  theme(plot.title = element_text(hjust = 0.5))
# secondary queue length versus time
qsl<-hat[hat$queue_label==5,]
qsl<-qsl[order(qsl$time),]
p4<-ggplot(qsl,aes(x=time,y=queue_length))+
  geom_step()+
  theme_bw()+ 
  theme(panel.border=element_blank(),panel.grid.major=element_blank(),panel.grid.minor=element_blank(),axis.line=element_line(colour = "black"))+
  labs(title="Secondary Queue Length")+
  ylab("Secondary Queue Length")+
  xlab("Time (minutes after simulation start)")+
  theme(plot.title = element_text(hjust = 0.5))
```

```{r arrival queue lengths by time,echo=FALSE,fig.cap='Length of the arrival queue as a function of the number of minutes after the start of the simulation. When a vehicle passes through, the arrival queue length momentarily jumps from 0 to 1 and back down to 0.'}
p0
```

```{r queue lengths by time,echo=FALSE,fig.cap='Primary and Secondary queue lengths as a function of the number of minutes fter the start of a simulation.'}
library(grid)
library(gridExtra)
grid.arrange(p1,p2,p3,p4,ncol=2)
```

```{r queue length summary,echo=FALSE,fig.cap='Fraction of time each queue spent with a specified queue length.'}
# percent of time a queue has length x
ptqx<-function(qol,x) {
  qol2<-rbind(c(0,NA,0),qol,c(max(qol$time),NA,NA))
  rm<-sum(qol2$time[which(qol2$queue_length==x)+1]-
           qol2$time[which(qol2$queue_length==x)])/max(qol2$time)
  return(rm)
}
# percent of time spent in each queue length state
ptqz<-function(qol) {
  sqs<-rep(NA,max(qol$queue_length)+1)
  for(i in 0:max(qol$queue_length)) sqs[i+1]<-ptqx(qol,i)
  return(sqs)
}
# dataframe for plotting
dffp<-function(hat) {
  ql<-NA
  ts<-NA
  qa<-NA
  qn<-c("Arrival","Lane 1","Lane 2","Lane 3","Secondary")
  for(i in 1:5) {
   qol<-hat[hat$queue_label==i,]
   qol<-qol[order(qol$time),]
   tis<-ptqz(qol)
   ts<-c(ts,tis)
   ql<-c(ql,c(0:(length(tis)-1)))
   qa<-c(qa,rep(qn[i],length(tis)))
  }
 return(data.frame(queue_length=factor(ql),time_in_state=ts,queue_label=qa)[-1,])
}
ppp<-dffp(hat)
ggplot(ppp)+
  aes(x=queue_label,y=time_in_state,fill=queue_length)+
  geom_col()+
  theme_bw()+ 
  theme(panel.border=element_blank(),panel.grid.major=element_blank(),panel.grid.minor=element_blank(),axis.line=element_line(colour = "black"))+
  labs(title="Queue Lengths")+
  ylab("Fraction of Time")+
  xlab("Queue")+
  theme(plot.title = element_text(hjust = 0.5))
```