---
title: "NINESIM Discrete Event Simulation Plots"
author: "Jan Irvahn"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r read in data,echo=FALSE}
system("python qdriver.py > out.txt")
system("grep \"\\[\" out.txt > timestamps.csv")
system("sed -i -e 's/\\[//g' timestamps.csv") 
system("sed -i -e 's/\\]//g' timestamps.csv")

st<-read.csv("timestamps.csv",header=FALSE)
names(st)<-c("arrival","leave_arrival","enter_primary_queue",
             "leave_primary_queue/start_primary_screening",
             "end_primary_screening_if_secondary_bound","enter_secondary_queue",
             "leave_secondary_queue/start_secondary_screening",
             "leave_secondary_screening_OR_leave_primary_for_exit",
             "exit")
primary_processing_time<-function(ol) {
  if(ol[5]==0) return(ol[8]-ol[4])
  if(ol[5]!=0) return(ol[5]-ol[4])
}
ppt<-apply(st,1,primary_processing_time)

line_mql<-system("grep \"QMAX\" qdriver.py",intern=TRUE)
snas<-unlist(strsplit(unlist(strsplit(line_mql,split="\\["))[2],","))
pl1qm<-as.numeric(snas[3])
pl2qm<-as.numeric(snas[4])
pl3qm<-as.numeric(snas[5])
sqm<-as.numeric(snas[6])

line_n<-system("grep \"NUM_VEHICLES\" qdriver.py",intern=TRUE)
n<-as.numeric(unlist(strsplit(unlist(strsplit(line_n,split=":"))[2],","))[1])

line_p<-system("grep \"probability of being sent to\" qdriver.py",intern=TRUE)
p<-as.numeric(unlist(strsplit(unlist(strsplit(line_p,split=":"))[2],","))[1])

line_mtba<-system("grep \"mean time between arrivals\" qdriver.py",intern=TRUE)
mean_arrival<-as.numeric(unlist(strsplit(unlist(strsplit(line_mtba,split=":"))[2],","))[1])

line_mtip<-system("grep \"mean time in primary\" qdriver.py",intern=TRUE)
mean_primary<-as.numeric(unlist(strsplit(unlist(strsplit(line_mtip,split=":"))[2],","))[1])

line_mtis<-system("grep \"mean time in secondary\" qdriver.py",intern=TRUE)
mean_secondary<-as.numeric(unlist(strsplit(unlist(strsplit(line_mtis,split=":"))[2],","))[1])

line_tt<-system("grep \"time to drive to a queue\" qdriver.py",intern=TRUE)
transit_time<-as.numeric(unlist(strsplit(unlist(strsplit(line_tt,split=":"))[2],"#"))[1])

options(scipen=1, digits=2)
```

# Introduction

PNNL has started writing python code for a discrete event simulation of the queueing model shown in Figure \ref{EG} using the SimianPie package that was developed at LANL. 

## Parameter Settings

The queueing model has 3 lanes at primary screening. The time between arrivals was exponentially distributed with a mean of `r mean_arrival` minutes. Primary screening times were exponentially distributed with a mean of `r mean_primary` minutes. Secondary screening times were exponentially distributed with a mean of `r mean_secondary` minutes. Transit times to get to the primary queues, the secondary queue, and the exit were `r transit_time` minutes. Maximum queue length for the three primary queues were `r pl1qm`, `r pl2qm`, and `r pl3qm`. The maximum queue length for secondary was `r sqm`. The probability of a vehicle being sent to secondary was `r p`.  The simulation described here ran `r signif(n,5)` vehicles. 

\begin{figure}[!h]
\includegraphics{eventgraph.pdf}
\caption{Event graph for NINESIM queueing work. \label{EG}}
\end{figure}

\clearpage

# Plots of Simulation Results

```{r arrival queue wait times,echo=FALSE,fig.cap='Time spent in the arrival queue as a function of arrival time after the start of the simulation.',fig.height=3.5}
plot(st$arrival,
     st$leave_arrival-st$arrival,main="Arrival Queue Wait Times vs. Arrival Time",
     ylab="Minutes",
     xlab="Arrival Time (Minutes)")
```

```{r hist of arrival queue wait times,echo=FALSE,fig.cap='Histogram of the times spent in the arrival queue.',fig.height=3.5}
hist(st$leave_arrival-st$arrival,main="Arrival Queue Wait Times",
     xlab="Minutes",breaks=50)
```

```{r hist of primary queue wait times,echo=FALSE,fig.cap='Histogram of the times spent in the queue for primary screening.'}
hist(round(st$`leave_primary_queue/start_primary_screening`-st$enter_primary_queue,10),
     main="Primary Queue Wait Times",xlab="Minutes",breaks=20)
```

```{r hist of primary screening times,echo=FALSE,fig.cap='Histogram of the times spent in primary screening. This includes time waiting for space in the secondary queue.'}
hist(ppt,main="Primary Screening Times",xlab="Minutes",breaks=20)
```

```{r hist of secondary queue wait times,echo=FALSE,fig.cap='Histogram of the times spent in the queue for secondary screening.'}
hist(st$`leave_secondary_queue/start_secondary_screening`[st$enter_secondary_queue!=0]-
       st$enter_secondary_queue[st$enter_secondary_queue!=0],
     main="Secondary Queue Wait Times",xlab="Minutes",breaks=20)
```

```{r hist of secondary screening times,echo=FALSE,fig.cap='Histogram of the times spent in secondary screening.'}
hm<-st$leave_secondary_screening_OR_leave_primary_for_exit[st$enter_secondary_queue!=0]-
  st$`leave_secondary_queue/start_secondary_screening`[st$enter_secondary_queue!=0]
hist(hm,main="Secondary Screening Times",xlab="Minutes",
     breaks=seq(from=0,to=ceiling(max(hm)),by=1))
```

```{r total time by arrival time,echo=FALSE,fig.cap='Total time in system as a function of arrival time.'}
library(ggplot2)

mdf<-data.frame(st)
secondary<-rep("no",nrow(mdf))
secondary[mdf$enter_secondary_queue!=0]<-"yes"
ndf<-cbind(mdf,secondary)
ggplot(ndf,aes(x=arrival,y=exit-arrival,color=secondary))+
  geom_point()+
  theme_bw()+ 
  theme(panel.border=element_blank(),panel.grid.major=element_blank(),panel.grid.minor=element_blank(),axis.line=element_line(colour = "black"))+
  labs(title="Total Time in System vs. Arrival Time")+
  ylab("Total Time in System (minutes)")+
  xlab("Arrival Time (minutes)")+
  theme(plot.title = element_text(hjust = 0.5))
```

```{r total time in system,echo=FALSE,fig.cap='Histogram of total time in system.'}
ggplot(ndf,aes(x=exit-arrival,color=secondary,fill=secondary))+
  geom_histogram(alpha=0.5)+
  theme_bw()+ 
  theme(panel.border=element_blank(),panel.grid.major=element_blank(),panel.grid.minor=element_blank(),axis.line=element_line(colour = "black"))+
  labs(title="Total Time in System")+
  xlab("Total Time in System (minutes)")+
  theme(plot.title = element_text(hjust = 0.5))
```